---
title: "DTC_5300_Q2_Analysis"
author: "Andrew Nalundasan"
date: "3/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
source('../AN_code/q2_code.R', local = FALSE)
```


## Why you are running the analyses you are running
## Analysis
### How the analyses answer the question being asked, and what the result is

## Carefully interpreting the results
## Results
### Presenting the results in an appealing way

## Acknowledging the assumptions you are making in each analysis, how plausible those assumptions are in the context of your data, and any evidence you can provide backing up those assumptions
## Assumptions


## After doing all analyses related to a given question, provide a generalized answer to the main questions
## Conclusion

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
